---
title: "Using R"
author: Orville D. Hombrebueno
date: today
format:
  html:
    embed-resources: true
editor: visual
toc: true
code-link: true
code-fold: true
cache: true
---

# The Data

```{r}
#| label: preliminaries
#| warning: false
#| message: false

# load packages
library(tidyverse) 
library(knitr)
library(gtsummary)
library(ggstatsplot)
library(downlit)

# load the data
data <- tibble(
  Response = c(rep("Yes", 20), rep("No", 20))
)
```

## Preference in Using R

This data is collected during the ...

The first 6 rows of the data is shown in @tbl-preference.

```{r}
#| label: tbl-preference
#| warning: false
#| message: false
#| tbl-cap: "The first 6 observations of the data set."

data |> # calling the data
  head() |> # getting the first six rows of the data
  kable() # creating a table
```

## Descriptive Statistics

The frequency distribution of the variable in the data set is shown in @tbl-summary.

```{r}
#| label: tbl-summary
#| tbl-cap: "Summary statistics of the data set."
#| message: false

data |> # calling the data
  tbl_summary() 
```

# Hypothesis Testing

::: panel-tabset
## Two-Sided

### Research Question

Do the participants have a 50/50 chance of preferring to use R?

### Hypothesis and Test Statistics

-   $H_0:p=50\%$; proportions of preferring to use R are equal.

-   $H_a:p\ne50\%$; proportions of preferring to use R are different.

-   Significance level, $\alpha=5\%$.

-   **chi-squared goodness of fit test**

-   Success-failure condition is met.

### Results and Interpretation

The pie chart and test statistics is shown on @fig-test. Based on the pie chart and the *p*-value being lesser than 5% (*p*-value = .001), we reject the null hypothesis. There is a strong evidence that the proportion of surviving is different among passengers. This may be influenced by certain factors like class, gender, or age?

```{r}
#| label: fig-test
#| fig-cap: Pie chart and test statistics.

ggpiestats( # creating the figure
  data = data, # data
  x = Response, # variable
  bf.message = FALSE, # Bayes Factor
  legend.title = "Use R?" # legend name
)
```

## One-Sided

### Research Question

Is the preference of using R greater than the expected preference proportion of 50%?

### Hypothesis and Test Statistics

-   $H_0:p=50\%$; the preference of using R is equal to the expected preference proportion of 50%.

-   $H_a:p>50\%$; the preference of using R is greater than the expected preference proportion of 50%.

-   Significance level, $\alpha=5\%$.

-   **one proportion Z-test**

- Success-failure condition is met.

### Results and Interpretation

```{r}
#| label: test-os

prop.test(
  x = 711, # number of successes
  n = 2201, # total number of trials (711 + 1490)
  p = 0.35, # expected preference proportion
  alternative = "greater" # other options: "two.sided" or "less"
)
```

The test statistics is shown above. Based on the results and the *p*-value being lesser than 5% (*p*-value = 0.004), we reject the null hypothesis. There is strong evidence that the survival rate is less than the expected survival rate of 35%.
:::
